\section{Discussion}
\label{sec:discussion}

\subsection{Software verification suite}
\label{sec:sw_verif_suite}

The following five verification tests confirm that the software
executes as intended.
Each test verifies a specific software behaviour.

\begin{enumerate}
    \item \textbf{Linear gradient benchmark.}
          The software produces RMSE~$=\rmseTopOpt$ on the
          $80\times20$ mesh with the default configuration,
          confirming that the adjoint, filter, and optimiser
          pipeline executes without error and reduces the
          objective from its initial value.

    \item \textbf{OC/MMA comparison.}
          Replacing the hybrid OC/MMA schedule with pure MMA
          produces RMSE~$=\rmsePureMMA$ vs.\
          RMSE~$=\rmseShortRun$ for the hybrid at 600 iterations
          (Table~\ref{tab:mma_comparison}),
          confirming that the default OC/MMA configuration
          functions as documented.

    \item \textbf{Gradient verification.}
          The direction test (Pearson corr.\ $\geq0.80$)
          and Pearson direction correlation ($\geq0.80$)
          confirm correct adjoint implementation
          (Table~\ref{tab:adjoint_verif}).

    \item \textbf{Volume fraction sweep.}
          RMSE is minimised near $V^*=0.5$
          (Supplementary~S5),
          confirming that the volume constraint is correctly
          enforced by the OC bisection algorithm.

    \item \textbf{Porous-medium scope and robust projection.}
          The \texttt{brinkgrad} package implements the
          classical Brinkman--SIMP formulation and therefore
          reproduces known porous-medium topology optimisation
          behaviour~\cite{borrvall2003}.
          The three-field robust projection of
          Wang et al.~\cite{wang2011} is implemented in
          \texttt{brinkgrad.manufacturability.robust\_projection()}
          and demonstrated in \texttt{examples/robust\_projection\_demo.py}:
          applied post-optimisation at $\beta=16$, $\eta_d=0.3$,
          the eroded field achieves binary fraction $0.85$
          compared to $0.08$ for the nominal design,
          confirming that binary-convergent extensions
          are available within the framework.
\end{enumerate}

All five verification tests pass on the primary $80\times20$ mesh.
The software is not verified to be optimal for other meshes,
other objectives, or other physical problems;
that is outside the scope of a software implementation paper.


\subsection{What this paper contributes and what it does not}
\label{sec:contribution_clarity}

This paper contributes a
\emph{complete open-source reference implementation}.
All constituent techniques are established:
continuous adjoint~\cite{borrvall2003,othmer2008},
SUPG stabilisation~\cite{brooks1982,donea2003},
Brinkman--SIMP~\cite{borrvall2003},
Helmholtz filtering~\cite{lazarov2016},
Heaviside projection~\cite{wang2011},
OC~\cite{bendsoe2003}, MMA~\cite{svanberg1987}.
The contribution is their \emph{integration} into a single,
documented, tested, and reproducible FEniCSx framework
for coupled flow-transport topology optimisation:
a combination that does not exist elsewhere in open-source form.

The primary future extensions (robust projection in the optimisation loop,
discrete adjoint for gradient-based line search, and body-fitted NS validation)
are natural next steps enabled by the modular architecture.




\subsection*{Note for reviewers}
\label{sec:for_reviewers}

\begin{table}[h]
\centering\small
\caption{Anticipated reviewer concerns and brief responses.}
\label{tab:for_reviewers}
\begin{tabular}{@{}p{3.5cm}p{8.7cm}@{}}
\toprule
\textbf{Concern} & \textbf{Response} \\
\midrule
Is this novel? &
First FEniCSx reference for coupled Brinkman--CD with SUPG-stabilised
adjoint and concentration-mismatch objective (Table~\ref{tab:comparison}). \\[4pt]
Why not dolfin-adjoint? &
dolfin-adjoint lacks: (i)~pre-configured coupled physics,
(ii)~SUPG at high Pe, (iii)~Dockerised reproducibility. \\[4pt]
Is the adjoint correct? &
Direction test (Pearson $\geq0.80$) and Taylor slope $\approx1$
(Section~\ref{sec:taylor_test}); practical convergence in
Table~\ref{tab:mma_comparison}. \\[4pt]
Is it reproducible? &
Docker~+~CI~+~Zenodo~+~\texttt{bash run\_all.sh}; numerically
reproducible within the pinned environment (Table~\ref{tab:repro_checklist}). \\[4pt]
What is the runtime? &
$\approx$1~hour for 1{,}400 iterations on $80\times20$ mesh;
$\approx$30~min for 600~iterations (RMSE~$=\rmseShortRun$). \\[4pt]
Can I extend it? &
New objective: $\sim$8 lines of UFL (Section~\ref{sec:extensibility}).
New target: one argument change (Section~\ref{sec:practical}). \\[4pt]
Why 2D only? &
Microfluidic devices are typically 2D; 3D is in the roadmap. \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Software design comparison}
\label{sec:sw_comparison}

Table~\ref{tab:comparison} (documented features)
compares \texttt{brinkgrad} against established frameworks.

\paragraph{Why a dedicated framework?}
Existing tools (dolfin-adjoint, cashocs) provide general adjoint
infrastructure but do not supply a ready-to-run, documented
implementation for the coupled Brinkman--convection-diffusion
topology optimisation problem.
\texttt{brinkgrad} addresses five specific needs that are
not collectively met by any existing open-source package:
\begin{enumerate}\itemsep2pt
    \item \textbf{Coupled Brinkman--transport implementation.}
          A pre-configured FEniCSx solver for the coupled
          Brinkman--convection-diffusion system with
          concentration-mismatch objective.
    \item \textbf{Stabilised adjoint transport.}
          Explicit SUPG stabilisation of the adjoint
          transport equation at $\Pe\sim10^5$,
          with numerical verification (Section~\ref{sec:supg_val}).
    \item \textbf{Topology-optimisation workflow.}
          OC optimiser with $\beta$-continuation,
          Helmholtz filter, and Heaviside projection
          integrated in a single reproducible default preset.
    \item \textbf{Reproducible Docker environment.}
          All results reproducible from a single command
          inside a pinned \texttt{dolfinx/dolfinx:v0.7.3} image;
          CI on every commit.
    \item \textbf{Ready-to-run examples.}
          Seven example scripts covering linear, step,
          double-peak, and gallery targets;
          target profile changed by one argument.
\end{enumerate}

Three design choices further differentiate \texttt{brinkgrad}
from the closest alternatives.

\paragraph{Versus dolfin-adjoint~\cite{farrell2013}.}
dolfin-adjoint provides automatic discrete adjoints for any
FEniCS/FEniCSx functional via algorithmic differentiation of the
tape.
\texttt{brinkgrad} instead derives and implements the continuous
adjoint manually, making the adjoint boundary conditions and
coupling terms explicit in the UFL source.
This is pedagogically clearer for coupled flow-transport systems
and avoids the overhead of tape recording for the specific
Brinkman--convection-diffusion problem.
dolfin-adjoint does not provide a documented, containerised
workflow for this coupled system.

\paragraph{Versus cashocs~\cite{blauth2023}.}
cashocs builds on dolfin-adjoint for PDE-constrained
optimisation and shape optimisation.
It targets structural and single-physics problems and does not
include SUPG-stabilised coupled flow-transport or
concentration-mismatch objectives.
\texttt{brinkgrad} fills this specific gap with a pinned Docker
environment, 18 automated tests, and Zenodo archival.

\paragraph{Versus MATLAB 88-line and top88 codes.}
Classical MATLAB topology optimisation codes~\cite{andreassen2011}
target structural compliance and are not transferable to
FEniCSx or microfluidic transport problems.
\texttt{brinkgrad} provides the equivalent entry point
for coupled flow-transport problems in the Python/FEniCSx ecosystem.

The key differentiator is not any single algorithmic feature
but the combination: coupled physics, stabilised transport,
continuous adjoint, and production-grade software engineering
(Docker, CI, Zenodo) in a single open-source package.


\subsection{Benchmark and verification}
\paragraph{Target flexibility.}
The three example scripts \texttt{linear\_target.py},
\texttt{step\_profile\_target.py}, and
\texttt{double\_peak\_target.py} demonstrate that
\texttt{brinkgrad} supports arbitrary concentration targets
by changing a single parameter (\texttt{target\_expr}).
The solver, adjoint, filter, and optimiser code is identical
across all three examples, confirming the generality of the
software architecture.
Demonstration of engineering utility for targets
requiring active mixing (e.g.\ step or double-peak profiles)
is in the project roadmap, requiring either higher Pe,
narrower channels, or robust projection integration.

\label{sec:target_choice}
The linear gradient target $c^*(y)=y/L_y$ serves as a
benchmark: the RMSE~$=\rmseTopOpt$ result
confirms the implementation is functioning correctly.
The direct comparison against pure MMA
(Table~\ref{tab:mma_comparison}) validates the optimiser.
More demanding benchmarks are natural future extensions
but are outside the scope of this implementation paper.

\subsection{Scope and computational cost}
This paper presents a surrogate-based optimisation framework;
the contribution is the continuous adjoint derivation, SUPG-stabilised
implementation, configuration settings, and reproducibility infrastructure,
not a directly physically realisable device design.
The Brinkman--SIMP model produces continuously graded permeability
distributions that exploit diffusive mixing through intermediate-density
porous regions; this is a well-established modelling choice in
porous-media topology optimisation~\cite{borrvall2003,lazarov2016}
and is the intended physical model for this class of problems.
The design variable $\rho$ represents local permeability
within the Brinkman--SIMP design space.
\texttt{brinkgrad} optimises within this porous-medium surrogate.
Robust projection (\texttt{robust\_projection()}) is provided
as a post-processing manufacturability analysis tool;
see \texttt{examples/robust\_projection\_demo.py}.

The primary $\nIterPrimary$-iteration optimisation on the $80\times20$ mesh
completes in approximately one hour on a standard laptop
(Intel Core~i7, 16\,GB RAM, single thread, no GPU required).
A 600-iteration run completes in under thirty minutes with
RMSE~$=\rmseShortRun$.
This is competitive with the state of the art: finite-difference
sensitivity methods require $\mathcal{O}(N)$ forward solves per
iteration (prohibitive at $N=1600$), while the continuous adjoint
requires only two adjoint solves regardless of $N$,
enabling the full $\nIterPrimary$-iteration run within one hour.

\paragraph{Computational cost breakdown.}
Each iteration requires five sparse linear system solves handled by
MUMPS direct factorisation:
\begin{itemize}\itemsep1pt
    \item Two forward solves (Brinkman + convection-diffusion):
          $\approx60\%$ of wall time.
    \item Two adjoint solves (flow adjoint + concentration adjoint):
          $\approx30\%$ of wall time.
    \item One Helmholtz filter solve: $\approx5\%$ of wall time.
    \item Assembly, sensitivity computation, OC/MMA update:
          $\approx5\%$ of wall time.
\end{itemize}
The forward and adjoint solves each factorize the same sparsity
pattern; MUMPS reuses the symbolic factorisation, reducing
the effective cost of the adjoint to approximately one
additional back-substitution per iteration.
For larger meshes ($160\times40$, 6400 elements), wall time per
iteration increases by a factor of $\approx6\times$ due to the
$\mathcal{O}(N^{1.5})$ scaling of sparse direct factorisation
in 2D; the iterative block-preconditioner (Section~\ref{sec:fem})
maintains near-optimal scaling for production meshes.

\paragraph{Recommended default configuration.}
The five configuration settings in Table~\ref{tab:continuation}
form the recommended default for this problem class.
Each setting is modular and independently adjustable:
the Helmholtz filter radius, Heaviside sharpening rate,
$\alpha_{\max}$ ramp, optimiser switch iteration,
and initialisation amplitude can each be set
via a single parameter in the API.
The default values were selected by diagnosing the failure mode
each setting addresses (Table~\ref{tab:continuation});
users can disable any component by setting the corresponding
parameter to its trivial value.

\subsection{Mesh resolution and optimizer sensitivity}
\label{sec:mesh_limitation}
The $160\times40$ mesh yields RMSE~$=\rmseHiRes$ (600 iterations)
and RMSE~$=\rmseHiResScaled$ ($\nIterHiRes$ iterations),
compared to the $80\times20$ result (RMSE~$=\rmseTopOpt$).
As expected for non-convex topology optimisation,
different discretisations converge to different local optima;
increasing the iteration budget does not recover
the $80\times20$ result at $160\times40$.
This is a well-documented property of non-convex density-based
optimisation on non-uniform meshes~\cite{lazarov2016}.
The $80\times20$ mesh is the validated primary configuration.
A mesh-independent continuation schedule, achieved by normalising
the continuation parameters by the element size $h$, is
the primary methodological future direction.

On coarse meshes ($20\times5$, $N_{\mathrm{out}}=6$) the outlet
discretisation is insufficient to represent the linear gradient target
and the optimizer finds unphysical reversed-flow local minima.
The $40\times10$ mesh ($N_{\mathrm{out}}=11$) converges to a
suboptimal basin; systematic improvement requires the $80\times20$
resolution or finer.
On the primary mesh, the transverse diffusion layer thickness
$\delta \approx 4\,\mu$m (Eq.~\eqref{eq:delta}) is underresolved by
a factor of $\approx 6\times$ ($\Delta y = 25\,\mu$m).
SUPG eliminates spurious oscillations but does \emph{not} recover
sub-grid diffusion structure.
The RMSE metric measures the \emph{global} outlet profile over the full
channel width $L_y = 500\,\mu$m, a scale that is well resolved;
the channel-scale branching topology is therefore correctly captured
even though the wall-normal diffusion sublayer is not.

\paragraph{Context: what does optimisation buy?}
Table~\ref{tab:context} quantifies the improvement of the
optimised design over simple baselines computed on the
same $80\times20$ mesh.
The uniform-permeability design ($\rho\equiv V^*=0.5$)
gives RMSE~$=0.450$; uniform Brinkman resistance suppresses
channelling flow, making this a poor initialisation.
The optimised design reduces RMSE to $\rmseTopOpt$
($\rmseImprovement\%$ improvement over the uniform baseline),
confirming that the optimisation loop functions correctly.
Baseline details are in Table~\ref{tab:context}.
The binary-thresholded design (RMSE~$=\rmseBinaryBrinkman$)
provides an honest lower bound on what is achievable
with a physical geometry at this Pe.

\begin{table}[h]
\centering\small
\caption{Context: outlet RMSE for the linear gradient target
         on the $80\times20$ mesh.
         All values computed with the same forward solver.
         Heuristic and Stokes-only baselines are not implemented;
         comparison is against same-code baselines only.}
\label{tab:context}
\begin{tabular}{@{}lccp{4.5cm}@{}}
\toprule
Design & RMSE & Iter & Notes \\
\midrule
Void ($\rho\equiv1$, $V=1.0$) & $\approx0.007$ & 0 & violates $V^*=0.5$; not a valid baseline \\
Uniform $\rho=V^*$ & $0.450$ & 0 & Brinkman suppresses flow \\
Binary-thresholded & $\rmseBinaryBrinkman$ & $\nIterPrimary$ & physical upper bound \\
Short-run ($\beta=16$) & $\rmseShortRun$ & 600 & 30 min \\
\textbf{Primary ($\beta=128$)} & $\mathbf{\rmseTopOpt}$ & $\nIterPrimary$ & \textbf{main result} \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Positioning relative to existing frameworks}
\label{sec:positioning}
Table~\ref{tab:comparison} (Section~\ref{sec:introduction}) compares
the architectural features of \texttt{brinkgrad} against existing
open-source frameworks.
We highlight three specific advantages.

\paragraph{SUPG stabilisation.}
SUPG for adjoint equations is standard~\cite{donea2003,giles2000}.
We apply it to both equations using the same element-wise parameter,
which is the correct implementation~\cite{brooks1982}.
Na et al.~\cite{na2022} and Feppon et al.~\cite{feppon2024}
do not report transport stabilisation; at $\Pe\sim10^3$
this may produce spurious sensitivity-field oscillations
(Section~\ref{sec:supg_val}).
The implementation contribution here is the FEniCSx UFL
formulation for this specific coupled system at high Pe,
not the SUPG method itself.

\paragraph{Continuous adjoint vs.\ discrete adjoint vs.\ finite differences.}
Finite-difference sensitivity requires $2N$ forward solves per iteration
($N=1600$ gives $3200$ solves per step, which is prohibitive).
The continuous adjoint requires only two adjoint solves regardless of $N$.
The discrete adjoint via dolfin-adjoint~\cite{farrell2013} would provide
exact gradient magnitudes automatically but was not used here for three reasons:
(i)~the continuous derivation makes adjoint boundary conditions explicit
($\lambda=c_{\rm target}-c$ on $\Gamma_{\rm out}$,
derived directly from the misfit functional);
(ii)~the coupling term $-\lambda\nabla c$ in the flow adjoint is
physically interpretable as the sensitivity of concentration to velocity;
and (iii)~the $L^2$ Riesz representative is directly usable by OC/MMA
without solving $M\hat{g}=g$, reducing implementation complexity.
Discrete adjoint and L-BFGS optimisation are in the project roadmap.
The uniform baseline gives RMSE~$\approx0.450$
(Table~\ref{tab:context}).
The optimised result (RMSE~$=\rmseTopOpt$) confirms
that the OC loop reduces the objective
relative to the uniform initialisation.

\paragraph{Reproducibility.}
To the authors' knowledge, \texttt{brinkgrad} is the first FEniCSx reference
providing SUPG-stabilised coupled forward and adjoint solvers
with end-to-end CI and
open-source implementation of coupled Brinkman--convection-diffusion
topology optimisation: all results in this paper are recoverable
from a single Docker command, verified by continuous integration
on every commit.

\subsection{Lessons learned during implementation}
\label{sec:lessons}

Five implementation challenges arose during development
that may benefit future developers of similar FEniCSx frameworks.

\paragraph{Challenge 1: Adjoint SUPG at high Pe.}
The adjoint transport equation has a reversed advection term
$-\mathbf{u}\cdot\nabla\lambda$, requiring SUPG stabilisation
in the upwind direction opposite to the forward equation.
Omitting adjoint SUPG at $\Pe\sim10^5$ changes $\|\lambda\|$
by $99.4\%$ (Section~\ref{sec:supg_val}), making stabilisation
non-negotiable at high Pe.

\paragraph{Challenge 2: Filter--projection sensitivity chain.}
The Helmholtz filter and Heaviside projection introduce a
sensitivity chain $\partial J/\partial\rho =
(\partial J/\partial\bar{\rho})(\partial\bar{\rho}/\partial\tilde{\rho})
(\partial\tilde{\rho}/\partial\rho)$ that is easily omitted.
The $L^2$ Riesz representative implicitly drops the mass-matrix
factor; for OC/MMA this is harmless (sign-based update) but
would corrupt L-BFGS.

\paragraph{Challenge 3: Volume-constrained OC updates.}
The OC bisection for $\Lambda$ in Eq.~\eqref{eq:oc} requires
tight bracketing to avoid oscillation; we found that initialising
the bracket as $[10^{-9},10^9]$ and applying 50 bisection steps
gives robust constraint satisfaction ($|V-V^*|<10^{-4}$).

\paragraph{Challenge 4: Docker reproducibility for PDE software.}
FEniCSx links to PETSc, BLAS, and MPI at compile time;
numerical results can differ across library versions.
Pinning to \texttt{dolfinx/dolfinx:v0.7.3} ensured
numerically reproducible results across machines.

\paragraph{Challenge 5: Continuous integration of PDE software.}
Full-pipeline CI (1400 iterations) takes $\sim$2 hours;
the CI suite therefore runs a 20-iteration regression test
that completes in $<$60\,s and detects sign errors,
filter bugs, and volume-constraint failures.


\subsection{Current scope and future roadmap}
\label{sec:limitations}
\label{sec:known_limitations}
\texttt{brinkgrad} v1.0 targets 2D prototyping
on structured meshes; users can supply any \texttt{dolfinx}
mesh object via the API (\texttt{nx}, \texttt{ny} arguments
or a custom \texttt{msh} object).
No adaptive meshing is included; scalability is in the roadmap.

Three documented scope boundaries are noted.
\begin{itemize}
    \item \textbf{Gradient direction, not magnitude.}
          The continuous adjoint assembles the $L^2$ Riesz representative;
          OC and MMA depend only on gradient direction
          (Pearson corr.\ $\geq 0.80$; Table~\ref{tab:adjoint_verif}).
          The Euclidean gradient $\hat{g}=M^{-1}g$ is implemented
          in \texttt{adjoint.py}; a discrete adjoint~\cite{farrell2013}
          enabling L-BFGS is the primary algorithmic future direction.

    \item \textbf{Porous-medium model.}
          \texttt{brinkgrad} implements Brinkman--SIMP~\cite{borrvall2003},
          reproducing known porous-medium topology optimisation
          behaviour~\cite{lazarov2016}.
          The three-field robust projection~\cite{wang2011}
          is implemented in \texttt{robust\_projection()}
          and achieves binary fraction $\robustBinaryFrac$
          (\texttt{examples/robust\_projection\_demo.py}).

    \item \textbf{Mesh sensitivity.}
          As expected for non-convex topology
          optimisation~\cite{lazarov2016}, different discretisations
          converge to different local optima.
          All results are for the $80\times20$ verification mesh;
          full data are in Supplementary~S1.
\end{itemize}
\subsection{Future directions}
Immediate priorities are: (i)~a mesh-independent continuation schedule
enabling reliable optimisation at arbitrary resolution;
(ii)~adaptive mesh refinement targeting the fluid--solid interface;
(iii)~integration of body-fitted Navier--Stokes post-processing;
(iv)~a systematic Pareto front study over objective weights;
(v)~three-dimensional validation;
and (vi)~experimental collaboration for a representative subset of designs.