\section{Introduction}
\label{sec:introduction}

Density-based topology optimisation for Stokes flow, introduced by
Borrvall and Petersson~\cite{borrvall2003}, has been extended to
conjugate heat transfer~\cite{alexandersen2014}, scalar transport
problems~\cite{makhija2015}, and unsteady flows~\cite{kreissl2011}.
The Brinkman--SIMP approach replaces solid regions with a high-permeability
porous medium, enabling gradient-based optimisation of fluid topology
without remeshing.
Microfluidic systems provide a natural benchmark application because the
small length scales of soft-lithography devices make manual re-design
expensive and simulation-driven optimisation increasingly accessible~\cite{whitesides2006}.

Despite growing interest, several computational gaps remain in the
open-source literature on coupled flow-transport topology optimisation.

\subsection*{Gaps addressed by this work}

\emph{Gap~1 --- Open-source implementation of coupled
flow-transport topology optimisation.}
Existing open-source topology optimisation codes for fluidic systems
target pure flow~\cite{borrvall2003} or global scalar
mixing~\cite{na2022}.
FEniCSx-based frameworks exist for structural problems~\cite{dokken2023}
and Stokes flow~\cite{lazarov2016}, but a documented, tested,
and reproducible open-source implementation for the coupled
Brinkman--convection-diffusion system with a boundary
concentration-mismatch objective has not been published.
The linear gradient benchmark is chosen for verifiability
as a sanity check; the framework's primary contribution
is the implementation methodology, not the benchmark RMSE.

\emph{Gap~2 --- High-P\'{e}clet transport in FEniCSx TO.}
Existing FEniCSx topology optimisation
implementations~\cite{dokken2023} operate at $\Pe\ll1$
and do not require transport stabilisation.
At microfluidic P\'{e}clet numbers ($\Pe\sim10^2$--$10^5$),
SUPG stabilisation of both forward and adjoint equations
is required~\cite{brooks1982,donea2003}; we document
this standard technique in the FEniCSx UFL context.

\emph{Gap~3 --- End-to-end reproducibility.}
Topology optimisation results are sensitive to implementation choices ---
element type, solver tolerance, continuation schedule, and
initialisation --- that are rarely reported in full.
Published microfluidic topology optimisation results~\cite{na2022,feppon2024}
are not accompanied by open-source code, making independent
verification and extension difficult.
We address this by releasing the complete pipeline as an installable
Python package with a Docker container and continuous-integration
workflow that reproduces every result from a single command.

Table~\ref{tab:comparison} positions \texttt{brinkgrad} relative to existing frameworks.

\begin{table}[h]
\centering\scriptsize
\caption{Architectural comparison of open-source topology optimisation
         frameworks for microfluidic and flow-transport problems.
         \checkmark~=~implemented; {---}~=~not reported.
         Direct quantitative RMSE comparison is not possible because
         published results use different domain geometries, boundary
         conditions, and objective definitions.}
\label{tab:comparison}
\begin{tabular}{@{}p{2.3cm}p{1.05cm}p{1.05cm}p{1.05cm}p{1.05cm}p{1.05cm}p{1.05cm}@{}}
\toprule
\textbf{Feature} & BP03 & Na22 & Fe24 & d-adj & cashocs & \textbf{ours} \\
 & \cite{borrvall2003} & \cite{na2022} & \cite{feppon2024} & \cite{farrell2013} & \cite{blauth2023} & \\
\midrule
\multicolumn{7}{@{}l}{\textit{Physics}} \\
Coupled flow-transport    & {---} & \checkmark & \checkmark & {---} & {---} & \checkmark \\
Conc.-mismatch objective  & {---} & \checkmark & {---}       & {---} & {---} & \checkmark \\
SUPG adjoint stabilisation& {---} & {---}       & {---}       & {---} & {---} & \checkmark \\
Continuous adjoint        & \checkmark & {---}  & \checkmark & {---} & \checkmark & \checkmark \\
Discrete adjoint          & {---} & {---}       & {---}       & \checkmark & \checkmark & future \\
\midrule
\multicolumn{7}{@{}l}{\textit{Software engineering}} \\
FEniCSx backend           & {---} & {---}       & {---}       & \checkmark & \checkmark & \checkmark \\
Docker image (pinned)     & {---} & {---}       & {---}       & {---} & {---} & \checkmark \\
CI / automated tests      & {---} & {---}       & {---}       & {---} & {---} & \checkmark \\
Zenodo archival (DOI)     & {---} & {---}       & {---}       & {---} & {---} & \checkmark \\
Jupyter examples          & {---} & {---}       & {---}       & {---} & \checkmark & \checkmark \\
Open source (MIT)         & partial & {---}     & {---}       & \checkmark & \checkmark & \checkmark \\
\bottomrule
\end{tabular}
\end{table}

\subsection*{Contributions}

All techniques used --- continuous adjoint, SUPG stabilisation,
Brinkman--SIMP, Helmholtz filtering, Heaviside projection, OC, MMA ---
are established methods from the literature.
The contribution of this paper is their integration into a complete,
documented, tested, and reproducible open-source FEniCSx framework
for coupled flow-transport topology optimisation.
Specifically:

\begin{enumerate}
    \item \textbf{Open-source implementation of coupled
          Brinkman--convection-diffusion topology optimisation in FEniCSx.}
          A complete, installable Python package implementing the
          coupled forward problem, continuous adjoint
          (including the coupling term $-\lambda\nabla c$
          and misfit-derived Dirichlet outlet condition),
          and sensitivity computation.
          Existing frameworks (dolfin-adjoint~\cite{farrell2013},
          cashocs~\cite{blauth2023}) provide general adjoint
          capabilities but lack: (i)~a pre-configured implementation
          of the coupled Brinkman--CD system with SUPG stabilisation
          at $\Pe\sim10^2$--$10^5$; (ii)~a default configuration
          with failure-mode documentation; (iii)~end-to-end
          Docker/CI reproducibility; and (iv)~a concentration-mismatch
          objective with explicit adjoint derivation
          (Table~\ref{tab:comparison}, Section~\ref{sec:adjoint}).

    \item \textbf{Documented implementation of stabilised
          forward and adjoint transport at $\Pe\sim10^2$--$10^5$.}
          The standard SUPG stabilisation~\cite{brooks1982,donea2003}
          is applied consistently to both forward and adjoint
          transport equations and verified by finite-difference
          Pearson direction correlation $\geq 0.80$
          confirming correct descent direction
          (Section~\ref{sec:verification}).

    \item \textbf{Reproducible software architecture.}
          Docker container, GitHub Actions CI on every commit,
          Zenodo archival, pytest suite, and Jupyter notebook.
          All results in this paper are reproducible from a single
          command in approximately one hour on a standard laptop
          (Section~\ref{sec:reproducibility}).

    \item \textbf{Verification and software-validation suite.}
          Gradient verification (Taylor remainder, direction test),
          mesh sensitivity study (four meshes),
          binary gap quantification (Supplementary~S5),
          and OC/MMA schedule verification
          (pure-MMA output RMSE~$=\rmsePureMMA$
          vs.\ hybrid RMSE~$=\rmseShortRun$,
          confirming the default configuration functions as intended;
          Table~\ref{tab:mma_comparison}).
\end{enumerate}

\paragraph{Nature of contribution.}
This paper is a \emph{software implementation paper}.
The contribution is not a new optimisation algorithm,
not a record-breaking RMSE, and not a manufacturable device.
The contribution is a documented, tested, and reproducible
open-source implementation of adjoint-based topology optimisation
for a coupled Brinkman--convection-diffusion system in FEniCSx ---
something that does not currently exist in the open-source literature.

\paragraph{Scope and limitations.}
\texttt{brinkgrad}\footnote{The name \texttt{brinkgrad} coincidentally
matches an educational automatic-differentiation library
by Karpathy~\cite{karpathy2020micrograd}.
This work is entirely independent, unrelated, and distinguished
by its scope: PDE-based topology optimisation of coupled
Brinkman--convection-diffusion systems.
The authors will adopt a unique name (e.g.\ \texttt{brinkgrad})
for the camera-ready version to avoid indexing ambiguity.}
operates within the Brinkman--SIMP porous-medium surrogate
on a fixed $80\times20$ mesh.
Two open directions are identified and quantified:
(i)~binary-constrained optimisation via robust projection~\cite{wang2011}
(thresholding results: Supplementary~S5); and
(ii)~mesh-independent continuation
(as expected for non-convex topology optimisation,
          finer meshes may find different local optima;
RMSE $0.091$--$0.107$ on $160\times40$
vs.\ $\rmseTopOpt$ on the $80\times20$ primary mesh).

The remainder of this paper is organised as follows.
Section~\ref{sec:model} presents the mathematical model and adjoint
derivation.
Section~\ref{sec:methods} describes the FEniCSx implementation and
configuration and continuation settings
(Section~\ref{sec:continuation}).
Section~\ref{sec:verification} presents verification studies.
Section~\ref{sec:results} presents the benchmark application.
Section~\ref{sec:discussion} discusses limitations and future directions,
and Section~\ref{sec:conclusion} concludes.