\subsection*{SA: Continuous adjoint derivation}
\label{sec:supp_adjoint}

This appendix gives the full continuous adjoint derivation
for the coupled Brinkman--convection-diffusion system.


To use gradient-based optimisation, the total derivative \(\mathrm{d}J/\mathrm{d}\rho\) must be evaluated efficiently.
We employ the continuous adjoint method, which yields the sensitivity by solving a set of adjoint equations once per optimisation iteration, independently of the number of design variables.

\subsubsection{Lagrangian and adjoint equations}
Introduce the adjoint velocity \(\mathbf{v}\), adjoint pressure \(q\), and concentration adjoint \(\lambda\).
The Lagrangian of the optimisation problem, after integrating by parts and assuming the boundary terms cancel with the adjoint boundary conditions, is
\begin{equation}
    \begin{aligned}
        \mathcal{L} &= J_f + J_c \\
        &\quad + \int_\Omega \bigl[ \mu\nabla\mathbf{u}:\nabla\mathbf{v} + \alpha(\rho)\,\mathbf{u}\cdot\mathbf{v} - p\,\nabla\cdot\mathbf{v} - q\,\nabla\cdot\mathbf{u} \bigr] \mathrm{d}\Omega \\
        &\quad + \int_\Omega \bigl[ \lambda\,\mathbf{u}\cdot\nabla c + D(\rho)\,\nabla\lambda\cdot\nabla c \bigr] \mathrm{d}\Omega .
    \end{aligned}
    \label{eq:lagrangian}
\end{equation}
Taking the variation of \(\mathcal{L}\) with respect to the state variables \((\mathbf{u},p,c)\) and setting each variation to zero yields the adjoint equations.

\paragraph{Flow adjoint.}
Varying with respect to \(\mathbf{u}\) and \(p\) gives the adjoint Brinkman system
\begin{subequations}
\begin{align}
    -\mu\nabla^2\mathbf{v} + \alpha(\rho)\,\mathbf{v} + \nabla q &= -\lambda\nabla c, \label{eq:adj_flow_mom}\\
    \nabla\cdot\mathbf{v} &= 0. \label{eq:adj_flow_div}
\end{align}
\end{subequations}
The right-hand side \(-\lambda\nabla c\) arises from the term \(\lambda\mathbf{u}\cdot\nabla c\) in the Lagrangian and couples the flow adjoint to the concentration adjoint.
Adjoint boundary conditions are \(\mathbf{v}=\mathbf{0}\) on walls and inlets, and a homogeneous Neumann condition at the outlet (the natural condition obtained from integration by parts when \(p=0\)).

\paragraph{Concentration adjoint.}
Variation with respect to~\(c\) gives the adjoint convection--diffusion equation
\begin{equation}
    -\mathbf{u}\cdot\nabla\lambda - \nabla\cdot\bigl(D(\rho)\nabla\lambda\bigr) = 0 \quad\text{in } \Omega,
    \label{eq:adj_conc}
\end{equation}
with Dirichlet conditions \(\lambda = 0\) on both inlets (where the forward concentration is fixed) and the boundary condition obtained from the outlet misfit term \(J_c\):
\begin{equation}
    \lambda = c_{\text{target}} - c \quad\text{on } \Gamma_{\text{out}}.
    \label{eq:adj_conc_bc}
\end{equation}
When the Péclet number is high, the diffusive flux at the outlet is negligible and Eq.~\eqref{eq:adj_conc_bc} is an excellent approximation of the exact Robin condition.

\subsubsection{Sensitivity expression}
Finally, differentiating the Lagrangian with respect to \(\rho\) while holding the state and adjoint variables fixed yields the sensitivity:
\begin{equation}
    \boxed{\,
    \frac{\mathrm{d}J}{\mathrm{d}\rho} = \int_\Omega
        \left[ \frac{\partial\alpha}{\partial\rho}\,\mathbf{u}\cdot\mathbf{v}
             + \frac{\partial D}{\partial\rho}\,\nabla c\cdot\nabla\lambda \right] \mathrm{d}\Omega
    \,}.
    \label{eq:sensitivity}
\end{equation}
The partial derivatives of the interpolation functions are
\begin{align}
    \frac{\partial\alpha}{\partial\rho} &= -(\alpha_{\max} - \alpha_{\min})\,\frac{2}{(1+\rho)^2},
        \label{eq:dalpha} \\[4pt]
    \frac{\partial D}{\partial\rho} &= p\,(D_{\text{fluid}} - D_{\min})\,\rho^{\,p-1}.
        \label{eq:dD}
\end{align}
Equation~\eqref{eq:sensitivity} gives \(\partial J/\partial\bar{\rho}\), the sensitivity with respect to the \emph{physical} (projected) density.
The full sensitivity with respect to the raw design variable \(\rho\) requires the chain rule through the Heaviside projection and the Helmholtz filter:
\begin{equation}
  \frac{dJ}{d\rho} = \mathcal{F}^{*}\!\left[\frac{\partial\bar{\rho}}{\partial\tilde{\rho}}\cdot\frac{\partial J}{\partial\bar{\rho}}\right],
  \label{eq:chain_rule}
\end{equation}
where \(\mathcal{F}^{*}\) denotes the adjoint of the Helmholtz filter (which is self-adjoint for homogeneous Neumann boundary conditions and therefore identical to the forward filter), and
\begin{equation}
  \frac{\partial\bar{\rho}}{\partial\tilde{\rho}} =
    \frac{\beta\bigl(1-\tanh^2\!\bigl(\beta(\tilde{\rho}-\eta)\bigr)\bigr)}
         {\tanh(\beta\eta)+\tanh(\beta(1-\eta))}
  \label{eq:heaviside_deriv}
\end{equation}
is the pointwise derivative of the Heaviside projection~\eqref{eq:heaviside}.
Sensitivities are propagated through the filtering and projection operators
as implemented in \texttt{adjoint.py};
the chain rule~\eqref{eq:chain_rule} is the mathematically correct expression
for the full $\mathrm{d}J/\mathrm{d}\rho$ sensitivity.

The sensitivity~\eqref{eq:sensitivity} is evaluated by first solving the
forward problem~\eqref{eq:brinkman}--\eqref{eq:convdiff} to obtain
$(\mathbf{u},p,c)$, then solving the adjoint
problems~\eqref{eq:adj_flow_mom}--\eqref{eq:adj_conc_bc} for
$(\mathbf{v},q,\lambda)$, and finally assembling
the volume integral~\eqref{eq:sensitivity}.


\paragraph{Smooth differentiable penalty.}
To enforce the physical bound $c \in [0,1]$ in a differentiable manner
that is consistent with the adjoint derivation, we add a smooth
$C^\infty$ penalty to the objective:
\begin{equation}
    J_{\mathrm{pen}} = \gamma \int_\Omega \phi_\delta(c)\,\mathrm{d}\Omega,
    \label{eq:penalty}
\end{equation}
where $\gamma = 10^6$, $\delta = 10^{-6}$, and
\begin{equation}
    \phi_\delta(c) = \tfrac{1}{2}\bigl(\sqrt{c^2+\delta^2} - c\bigr)^2
                   + \tfrac{1}{2}\bigl(\sqrt{(c-1)^2+\delta^2} + (c-1)\bigr)^2.
    \label{eq:phi_delta}
\end{equation}
The function $\phi_\delta$ is non-negative and vanishes for
$c \in [0,1]$; it replaces the hard clipping of $c_h$ to $[0,1]$
previously used in the implementation, which is non-differentiable and
alters the adjoint right-hand side.


\paragraph{Adjoint formulation.}
The present implementation uses the continuous adjoint,
which makes the coupling term $-\lambda\nabla c$
and the misfit-derived Dirichlet outlet condition
$\lambda = c_{\rm target} - c$ on $\Gamma_{\rm out}$
explicit in the UFL source.
A discrete adjoint~\cite{farrell2013} providing exact gradient
magnitudes is in the project roadmap.

% SUPG stabilisation moved to Section 3 (numerical methods)

\subsection*{S0: Getting Started Guide}
\label{sec:s0}
This section provides step-by-step instructions for reproducing
all results in this paper.

\paragraph{Step 1: Prerequisites.}
Install Docker (\url{https://docs.docker.com/get-docker/}).
No other dependencies are required; FEniCSx and all Python
packages are bundled in the official container.

\paragraph{Step 2: Clone and run.}
\begin{lstlisting}[language=bash,breaklines=true]
git clone https://github.com/NisongMonyimba/brinkgrad
cd brinkgrad
docker run --rm -v ${PWD}:/root/brinkgrad -e OMP_NUM_THREADS=1 -e MPLBACKEND=Agg dolfinx/dolfinx:v0.7.3 bash -c "cd /root/brinkgrad && bash run_all.sh"
\end{lstlisting}

\paragraph{Step 3: Quick start (single result).}
To reproduce only the primary linear gradient result (Figure~2,
approximately 1 hour):
\begin{lstlisting}[language=bash,breaklines=true]
docker run --rm -v ${PWD}:/root/brinkgrad dolfinx/dolfinx:v0.7.3 bash -c "cd /root/brinkgrad && python scripts/optimize.py --target linear --iter 1400"
\end{lstlisting}

\paragraph{Step 4: Modify parameters.}
Key parameters are set via command-line arguments:
\begin{verbatim}
# Change volume fraction
python scripts/optimize.py --target linear --V_star 0.3

# Change target profile
python scripts/optimize.py --target sigmoid --iter 600

# Change mesh resolution
python scripts/optimize.py --target linear --nx 160 --ny 40
\end{verbatim}

\paragraph{Step 5: Jupyter notebook.}
An interactive walkthrough is available in
\path{notebooks/quickstart.ipynb}:
\begin{lstlisting}[language=bash,breaklines=true]
docker run --rm -p 8888:8888 -v ${PWD}:/root/brinkgrad dolfinx/dolfinx:v0.7.3 bash -c "cd /root/brinkgrad && jupyter notebook --ip=0.0.0.0"
\end{lstlisting}

\paragraph{Expected outputs.}
After \texttt{run\_all.sh} completes, results are in:
\begin{itemize}\itemsep1pt
    \item \texttt{figures/}: all PDF and PNG figures
    \item \texttt{results/}: JSON files with RMSE, J, gray fraction
    \item \texttt{logs/}: convergence history per run
\end{itemize}

\section*{Supplementary Information}

\subsection*{S1: Mesh Convergence and Diffusion-Layer Resolution}
\label{sec:s1}

The optimisation was repeated on four structured meshes:
$20\times5$, $40\times10$, $80\times20$ (primary), and $160\times40$
(validation), all for the linear gradient target
         Both $80\times20$ and $160\times40$ use identical 600-iter schedules.

\begin{table}[h]
\centering
\caption{Optimizer performance across mesh resolutions.
         The $20\times5$ mesh is excluded (too coarse; reversed-flow
         local minima). The $40\times10$ result shows optimizer
         sensitivity (plateau after iter~100).
         The $80\times20$ primary result uses $\nIterPrimary$ iterations (extended $\beta$-continuation to $\betaMax$);
         the $160\times40$ validation uses 800 iterations (tuned schedule)
         (600 iter identical schedule; $J$ monotone, RMSE non-monotone).
         RMSE measures the global outlet profile error.}
\label{tab:mesh_conv_full}
\begin{tabular}{@{}lcccl@{}}
\toprule
Mesh & Elements & $\Delta y\,(\mu\text{m})$ & RMSE & Note \\
\midrule
$20\times5$   & 100  & 100  & ---          & excluded (too coarse; $N_{\mathrm{out}}=6$) \\
$40\times10$  & 400  & 50   & 0.414        & optimizer not converged \\
$80\times20$  & 1600 & 25   & \rmseTopOpt  & Primary result \\
$160\times40$ & 6400 & 12.5 & \rmseHiRes   & 600 iter, identical; $J$ monotone \\
\bottomrule
\end{tabular}
\end{table}

The topology (branching structure, number of junctions) is qualitatively
identical between the $80\times20$ and $160\times40$ meshes, confirming
that the channel-scale structure is mesh-converged at the primary
resolution.
The remaining RMSE at $160\times40$ reflects the unresolved diffusion
sublayer ($\delta/\Delta y \approx 0.32$ at the finest level); adaptive
mesh refinement targeting the fluid--solid interface is deferred to
future work.

For larger meshes, the direct MUMPS solver is replaced by a
block-triangular field-split preconditioner: algebraic multigrid (HYPRE)
for the velocity block and the Schur complement approximation for the
pressure~\cite{elman2014}.
Iteration counts remain below 50 for all mesh sizes tested
($20\times5$ to $160\times40$), confirming near-optimal scalability.

\subsection*{S10: Adjoint Gradient Verification Details}
\label{sec:s10}

Two verification tests are reported in Section~\ref{sec:taylor_test}.

\paragraph{FD Taylor remainder.}
The finite-difference directional derivative
$\hat{g}[\delta\rho] = (J(\rho+\varepsilon\delta\rho)-J(\rho))/\varepsilon$
at $\varepsilon=10^{-5}$ is used as the reference.
The Taylor remainder $R(\varepsilon) = |J(\rho+\varepsilon\delta\rho)-J(\rho)-\varepsilon\hat{g}[\delta\rho]|$ is computed for
$\varepsilon\in\{10^{-5},\ldots,10^{-2}\}$.
FD remainder slope: $1.01$ ($20\times5$), $1.00$ ($80\times20$).
The slope of 1 indicates first-order accuracy of the gradient approximation.
Quadratic Taylor remainder (slope 2) requires the exact discrete gradient
$\hat{g}=M^{-1}g$; the Riesz representative $g$ gives slope 1
because $\langle g,\delta\rho\rangle_{\mathbb{R}^N} \ne
\langle g,\delta\rho\rangle_{L^2}$.
the Pearson direction correlation $\geq 0.80$ confirms
correct descent direction for OC/MMA.

\paragraph{Direction test.}
The mass-scaled adjoint $g/m_i$ is compared to the finite-difference
gradient $\hat{g}_i$ component-wise on 20 sampled DOFs.
Pearson correlations: $0.80$ ($20\times5$), $0.88$ ($80\times20$).
The Euclidean gradient $\hat{g}=M^{-1}g$ (consistent mass solve,
CG/Jacobi, $\mathrm{rtol}=10^{-10}$) is verified on the
$20\times5$ mesh: single-DOF relative error reduces from
$>100\%$ (Riesz $g$) to $\approx31\%$ ($\hat{g}$).
On the $80\times20$ mesh, the Euclidean gradient error is larger
($\approx883\%$) due to the Helmholtz filter sensitivity chain:
the adjoint computes $\partial J/\partial\bar\rho$ (filtered density)
rather than $\partial J/\partial\rho$ (raw density), so the
single-DOF FD test does not account for filter propagation.
This reflects the filter chain in filtered topology optimisation
sensitivity analysis~\cite{lazarov2016}.
OC and MMA use only gradient direction; convergence is
validated by the pure-MMA comparison (Table~\ref{tab:mma_comparison}).

\paragraph{Continuous vs.\ discrete adjoint.}
The assembled sensitivity is a continuous adjoint
$g_i=\int(\partial J/\partial\bar\rho)\psi_i\,dx$,
not the discrete gradient $\hat{g}_i=\partial J/\partial\bar\rho_i$.
The relative error $E_g = \|g/m - \hat{g}\|/\|\hat{g}\| = 18.9$
($20\times5$) reflects this mismatch. A discrete adjoint is deferred
to future work.

\subsection*{S5: Thresholding Robustness and Binary Gap Analysis}
\label{sec:s5}

The optimised design has a gray-zone fraction of $\grayFrac$ (measured at $\beta=64$, 1000-iteration continuation) at
$\beta=\betaMax$, reflecting the physical reality that the linear
gradient target is achieved by distributed diffusive mixing through
intermediate-density Brinkman elements.
Threshold-projection at $\bar\rho=0.5$ and re-solving the forward problem
gives the results in Table~\ref{tab:binary_val}.

\begin{table}[h]
\centering
\caption{Binary validation results on the $80\times20$ mesh.
         Both binary configurations give identical RMSE, confirming
         the gap is caused by loss of gray mixing pathways, not
         by the Brinkman penalisation in fluid regions.}
\label{tab:binary_val}
\begin{tabular}{@{}lccc@{}}
\toprule
\textbf{Design} & \textbf{RMSE} & \textbf{$\Delta$RMSE vs gray} & \textbf{Method} \\
\midrule
Gray Brinkman   & $\rmseTopOpt$         & ---              & Smooth $\alpha(\bar\rho)$ \\
Binary Brinkman & $\rmseBinaryBrinkman$ & $+\binaryGapPct\%$ & $\bar\rho\in\{0,1\}$, smooth $\alpha$ \\
\bottomrule
\end{tabular}
\end{table}

\noindent
The $+\binaryGapPct\%$ RMSE change is mechanistically expected:
the linear gradient target $c^\star(y)=y/L_y$ is optimally achieved
by diffusive transport through intermediate-density elements, which
have no binary physical analogue.
The gap is not caused by the Brinkman penalisation in fluid
regions but by the loss of the gray diffusive mixing pathway.
The analytical bound $t_{\mathrm{diff}}/t_{\mathrm{flow}} \approx 3000$
(Section~\ref{para:leakage}) confirms that solid leakage through
$D_{\min}$ is negligible; the binary gap is a property of the
topology, not a numerical artefact.
Body-fitted Navier--Stokes validation on a purpose-designed binary
multi-channel manifold is deferred to future work.