\section{Conclusion}
\label{sec:conclusion}

We have presented \texttt{brinkgrad}, a complete open-source reference
implementation integrating established techniques into a reproducible
FEniCSx framework
for density-based topology optimisation of coupled
Brinkman--convection-diffusion systems.
The implementation makes four documented contributions:

\begin{enumerate}
    \item \textbf{Continuous adjoint implementation.}
          A complete FEniCSx implementation of the continuous adjoint
          for coupled Brinkman--convection-diffusion, following
          standard Lagrangian methodology~\cite{borrvall2003},
          with the momentum--concentration coupling $-\lambda\nabla c$
          and misfit-derived Dirichlet outlet condition.

    \item \textbf{FEniCSx UFL implementation including SUPG.}
          Standard SUPG~\cite{donea2003} applied to both forward
          and adjoint transport equations, verified to reduce
          sensitivity-field oscillations from $\sim0.4$ to
          below $10^{-3}$ at $\Pe\sim10^2$--$10^5$.

    \item \textbf{Documented default configuration.}
          Five modular settings with failure-mode motivation
          (Table~\ref{tab:continuation});
          reduces gray fraction from $\grayShortRun$ to $\grayFrac$.
          The hybrid OC/MMA split is validated by direct comparison
          against pure MMA (Table~\ref{tab:mma_comparison}).

    \item \textbf{Open-source reproducible release.}
          MIT licence, Docker, CI, and Zenodo archival;
          all results reproducible from a single command
          in approximately one hour on a standard laptop.
\end{enumerate}

As a benchmark application, the framework is demonstrated on
concentration gradient generation in porous-medium flow.
On an $80\times20$ mesh, the framework achieves
RMSE~$=\rmseTopOpt$ for the linear gradient benchmark,
confirming correct execution of the adjoint--optimiser pipeline
(gray-zone fraction $\grayFrac$).
\texttt{brinkgrad} optimises within the Brinkman--SIMP
porous-medium model; gray-region behaviour is characteristic
of density methods and is documented in Supplementary~S5.
The three-field robust projection~\cite{wang2011} is implemented
in \texttt{brinkgrad.manufacturability.robust\_projection()}
and demonstrated in \texttt{examples/robust\_projection\_demo.py};
integration into the optimisation loop is the primary future direction.

\paragraph{Future work.}
The principal open directions are:
(i)~robust projection~\cite{wang2011} enforcing near-binary designs
during optimisation --- the primary open direction;
(ii)~a discrete adjoint for exact gradient magnitude;
(iii)~three-dimensional extension;
(iv)~adaptive mesh refinement targeting the diffusion sublayer
($\delta\approx 4\,\mu$m);
(v)~experimental validation on a binary-projected design.

The framework is designed to be general: any differentiable functional
of velocity and concentration fields can be incorporated by modifying
the objective and adjoint boundary conditions in the UFL description.
This generality, combined with a computational cost of under one hour on a standard laptop
(or under thirty minutes for the 600-iteration quick run)
and full end-to-end reproducibility,
makes \texttt{brinkgrad} a practical foundation for surrogate-based
computational flow-transport optimisation and a reference implementation
for coupled flow-transport adjoint methods in FEniCSx.